---
title: "COVID-19 Testing Trends [Guided Project]"
author: "Jonathan"
date: "12/22/2020"
output: html_document
---

## Summary
Using a dataset provided by Kaggle, this project aims to build my skills and understanding of the data analysis workflow by evaluating the COVID-19 situation.
Source: https://www.kaggle.com/lin0li/covid19testing?select=tested_worldwide.csv
Note: All N/A values are replaced with 0 for this test.
  Due to the lack of observations and recordings, this test is deemed to be inconclusive.
  

## Questions
Which country has the highest number of positive cases against the number of tests?

## Code
``` {r part1}
# read and setup the dataframes
# record some general information
library(readr)
library(tidyverse)
covid_df <- read_csv("tested_worldwide.csv")
covid_df[is.na(covid_df)] <- 0
dimension <- dim(covid_df)
vector_cols <- colnames(covid_df)

head(covid_df)
glimpse(covid_df)
```

```{r part2}
# filter the dataframe as we do not consider state specific cases
covid_df_all_states <- covid_df %>% filter(Province_State == "All States") %>% select(-Province_State)
covid_df <- covid_df %>% select(-Province_State)
```

```{r part3}
# select important daily information
covid_df_all_states_daily <- covid_df_all_states %>%
  select(Date, Country_Region, active, hospitalizedCurr, daily_tested, daily_positive)
```

```{r part4}
# find the sum of useful info and retrieve the top 10s
covid_df_all_states_daily_sum <- covid_df_all_states_daily %>% group_by(Country_Region) %>% summarize(
  tested = sum(daily_tested), positive = sum(daily_positive),
  active = sum(active), hospitalized = sum(hospitalizedCurr)) %>% arrange(-tested)

covid_top_10 <- head(covid_df_all_states_daily_sum, 10)
```

```{r part5}
# Split the information into different vectors, find the ratio, select top 3 
countries <- covid_top_10$Country_Region

tested_cases <- covid_top_10$tested
names(tested_cases) <- countries

positive_cases <- covid_top_10$positive
names(positive_cases) <- countries

active_cases <- covid_top_10$active
names(active_cases) <- countries

hospitalized_cases <- covid_top_10$hospitalized
names(hospitalized_cases) <- countries

positive_tested_top_10 <- positive_cases / tested_cases

positive_tested_top_3 <- sort(positive_tested_top_10, decreasing = T)[1:3]
positive_tested_top_3

```

``` {r part6}
# Knowing top 3, create a presentable table
## Altered code so this works for any given data
country_names <- names(positive_tested_top_3)

first_pos <- positive_tested_top_3[country_names[1]]
first_stats <- covid_df_all_states_daily_sum %>% filter(Country_Region == country_names[1]) %>% select(-Country_Region)
first <- c(first_pos, first_stats)

second_pos <- positive_tested_top_3[country_names[2]]
second_stats <- covid_df_all_states_daily_sum %>% filter(Country_Region == country_names[2]) %>% select(-Country_Region)
second <- c(second_pos, second_stats)

third_pos <- positive_tested_top_3[country_names[3]]
third_stats <- covid_df_all_states_daily_sum %>% filter(Country_Region == country_names[3]) %>% select(-Country_Region)
third <- c(third_pos, third_stats)

covid_mat <- rbind(first, second, third)
colnames(covid_mat) <- c("ratio", "tested", "positive", "active", "hospitalized")
rownames(covid_mat) <- country_names
covid_mat
```

```{r part7}
# Test conclusions and answers to original question
question <- "Which countries have had the highest number of positive cases against the number of tests?"

answer <- c("Positive tested cases: " = positive_tested_top_3)
result_list_dataframes <- list(covid_df, covid_df_all_states, covid_df_all_states_daily, covid_df_all_states_daily_sum)
result_list_matrices <- list(covid_mat)
result_list_vectors <- list(vector_cols, countries)
data_structure_list <- list(dataframes = result_list_dataframes, matrices = result_list_matrices, vectors = result_list_vectors)

covid_analysis_list <- list(Question = question, "Answers & Results" = answer, "Data Structures" = data_structure_list)
covid_analysis_list[1:2]

```


The Sample List of Data Structures created in this program

``` {r final}
covid_analysis_list
```





